Geometric variational problems

### 2020-11-14 NSF2020/Description/background.tex
fixed typos in background

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
FSU), and Y.-N. Young (PI, NJIT)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent molecules (water).
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The substantial free energy for placing hydrophobic substances in contact with water 
is roughly proportional to the surface area of the contact region \cite{Bjorneholm2016}.  
%As a result, hydrocarbon solutes have a large interfacial tension 
%and try to minimize their surface area when in water. 
At the microscopic level, the hydrophobic force is a long-range, surface
interaction. This means that two hydrophobic surfaces, separated by
water over some distance, experience an attractive force
\cite{Lum1999,Meyer2006,Hammer2010}. Measurements show that the
hydrophobic force decays exponentially with a decay length like 1 nm
\cite{Israelachvili1984, Marcelja1977,Christenson2001,Lin2005}. It is
also known that the interaction is not pairwise additive, meaning that
the force between any two hydrophobic objects is altered by the presence
of a third object, hydrophobic or otherwise \cite{SilveraBatista1242477}.

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection $P_1,$ $P_2,$ $P_3, \ldots$ of rigid particles. The
  exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM} that is based
on the physical origin of hydrophobicity. This model addresses the major
shortcomings of molecular dynamics (MD) and continuum approaches. Based
on preliminary results (\S\ref{sec:preliminary_work}) the PIs propose to
extend this HAP model to offer an alternative modeling methodology that
leads to new mathematical ideas, and is both physically accurate and
computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's earlier work on
the problem of energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a mathematical, squared-gradient
theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators initiated a collaboration to
create a gradient theory for arbitrary collections of hydrophobic and
amphiphilic particles. This new method eliminates the costly calculation
of water by treating the solvent implicitly, is capable of representing
biologically relevant morphologies, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

@@

The above considerations motivate the following definition for the HAP:
\begin{align}
\label{HAP}
  \Phi = \gamma \int_{\Omega} \rho |\nabla u|^2 + \rho^{-1}u^2 \,\dif V. 
\end{align}
The integrand contains
the decay length $\rho$
and a dimensionless scalar function $u(\mathbf{x})$ called the water activity.
The integrand has units of an inverse length and so
the volume integral has units of area. Multiplication by interfacial tension
$\gamma$ makes $\Phi$ an energy.

To establish an attraction between interfaces, the water activity is not
arbitrary, but rather is the solution of the screened Laplace
equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0, \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$, describes a hydrophobic interface and
$f=0$ describes a hydrophilic interface. Thus $f$ encodes information
about the particles and the parameters $\rho$ and $\gamma$ encode
information about the quality of the solvent~\cite{Israelachvili1954,
Discher2002}. In practice, we integrate~\eqref{HAP} by parts and
use~\eqref{SL} to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding volume integral calculations.

The equations~\eqref{HAP} and~\eqref{SL} possess a number of the
mathematical properties that mirror the phenomenological characteristics
of hydrophobic attraction. Specifically, solutions of~\eqref{SL} yield
an attractive force between hydrophobic bodies separated by water, and
this attraction decreases exponentially with the distance of
separation~\cite{Eriksson1989}. Using boundary layer analysis, the HAP
converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}. Finally, we have
demonstrated that the forces derived from HAP theory are
non-additive~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics, let $P_1$, $P_2$, \ldots, $P_N$, be a
finite collection of disjoint, rigid, and closed particles with
Lipschitz boundary. Assume that the label $f$ is an element of the

@@

overlap. The total potential, force and torque are then

@@

To supply viscous dissipation, we incorporate the mobility problem
flow for a rigid body suspension in Stokes flow:
\begin{equation}
\label{eq:stokes}
\begin{aligned}
  &-\mu \Delta \mathbf{u} + \nabla p = 0, \quad \mathbf{x} \in \Omega, \qquad 
  \nabla \cdot \mathbf{u} = 0,  \quad \mathbf{x} \in \Omega,\\
  &{\bf u}(\mathbf{x}) \to 0 \quad \text{as}\ |\mathbf{x}|\to \infty,\qquad 
  \mathbf{u}(\mathbf{x})|_{\partial P_i} = \mathbf{v}_i +
\boldsymbol{\omega}_i\times(\mathbf{x} - \mathbf{a}_i),\\
&\int_{\partial P_i}\boldsymbol{\sigma}\cdot {\bf n} \dif S=-{\bf F}_i, \quad
\int_{\partial P_i} (\mathbf{x} - \mathbf{a}_i)\times (\boldsymbol{\sigma} \cdot \mathbf{n}) \dif S =-{\bf G}_i.
\end{aligned}
\end{equation}
Here $\mu$ is the fluid viscosity and the first two equations state that
the fluid motion is a divergence-free Stokes flow; the third equation
specifies that the fluid velocity vanishes at infinity; the fourth
equation enforces a rigid body motion on each particle, where
$\mathbf{v}_i$ and $\boldsymbol{\omega}_i$ are unknown translation and
angular velocities; the last two equations state that the net
viscous forces and torques with fluid shear stress $\boldsymbol{\sigma}$
balance \eqref{eq:total_poten}.

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

@@

It is worth emphasizing that the HAP model uses only a few parameters;
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrates our interfacial tension
parameter. This is in direct contrast with pair-potential based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of the
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparisons with experimental results \S\ref{subsec:specific_aim_1}, to
establish that the HAP model has the capability and adaptivity to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2}, using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15}, and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. These vesicle studies built off of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles, and red blood cells. These approaches range from
the finite element method~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15},
phase field method~\cite{Du05,QiangDu08,Lowengrub13}, and immersed
boundary method~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators
led in part the development of phase field functionals of membrane
elastic energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from these prior methods in a number of
respects. First, we do not assume a surface. Rather, we more
fundamentally consider a collection of amphiphilic particles. The
collection of amphiphiles minimize hydrophobic interactions by
sequestering hydrophobic tails in the form of a bilayer, and the
particles' excess free energy gives rise to an elastic bilayer energy.
The second difference lies in the fluid-interface coupling. Here, the
associated mobility problem~\eqref{eq:stokes} is more complicated than
dealing with a stress boundary condition or diffusive surface force,
because the fluid velocity must be that of a rigid body motion at the
particle surfaces. Finally, the HAP model directly addresses the
existence of multiple phases. We can vary lipid length, spontaneous
curvature, and bending rigidity by introducing different particle shapes
and hydrophobic boundary conditions (Figure~\ref{fig:demixing}). In
contrast, continuum theory deals with multiple phases through additional
surface densities that must satisfy specialized transport
equations~\cite{Lowengrub07, MikuckiZhou17}.

\begin{wrapfigure}[6]{r}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/Lysis.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:lysis} \footnotesize Two hydrophobic particles
  enter and then lyse the circular bilayer.}
\end{wrapfigure}
The greatest strength of using HAP to model lipid bilayer membrane is
the ability to form discontinuities (interfacial singularity) from first
physical principles without any artificial manipulation to rearrange the
interface (Figures~\ref{fig:demixing} and~\ref{fig:lysis}). For
example, using the HAP model to simulate a vesicle under a shear flow
that is strong enough to cause the vesicle membrane to rupture (see
\S\ref{sec:preliminary_work}), we expect the HAP model to capture the
reorganization of lipid molecules on the scales of membrane thickness
($\sim 5$ nm), which is a nearly impossible task using phase-field or
immersed boundary methods without extreme refinement around the membrane
and artificial treatment of reconnection during the topological change
of an interface \cite{doi:10.1063/5.0009734, LiAn-Chang16,
doi:10.1098/rspa.2012.0505, doi:10.1137/130941432, Feetzl18,
doi:10.1137/16M1108406}.

>>>>

\noindent
{\bf Collaborative Research: Mathematical modeling and simulation of
self-assembling amphiphilic particles in solvent} \\
%{\bf {Collaborative Research: Mathematical modeling and coarse-grained simulations of self-assembly of amphiphilic Janus particles in a solvent}}\\
{\em Rolf Ryham (lead PI, Fordham University), Bryan Quaife (PI,
Florida State University), and Young-Nan Young (PI, New Jersey Institute of Technology)}
\section{Background}
\label{sec:background}
The goal of this collaborative proposal is to use mathematical modeling
and numerical simulations to investigate the dynamic self-assembly of
amphiphilic particles interacting via hydrophobic forces
in a solvent. Amphiphilic particles (such as lipid molecules) possess
both hydrophobic and hydrophilic structures. In a viscous solvent they
self-assemble into meso-/macroscopic structures (such as micelles and
bilayers of lipids) to shield their hydrophobic parts from contact with
the solvent (water) molecules.
%
%\subsection{Hydrophobic Forces}
%\label{sec:hydrophobicforce}
%
Such self-assembly of amphiphiles via hydrophobic forces is ubiquitous in biology and biophysics \cite{Israelachvili1954},
%The hydrophobic force is a ubiquitous molecular interaction in biology \cite{Israelachvili1954}, 
and has been a major source of nonspecific interactions between
nanoparticles in soft matter
\cite{Sanchez-IglesiasEtAl2012_ACSNano,AltantzisEtAl2013_PSC,XieYangLuEtAl2020_COCIS}. 
%The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles so to design smart materials of desirable properties by
%tuning the geometry and properties of the amphiphilic particles.

@@

The substantial free energy for placing hydrophobic substances in contact with water 
is roughly proportional to the surface area of the contact region \cite{Bjorneholm2016}.  
%As a result, hydrocarbon solutes have a large interfacial tension 
%and try to minimize their surface area when in water. 
At the microscopic level, the hydrophobic force is a long-range, surface
interaction. This means that two hydrophobic surfaces, separated by
water over some distance, experience an attractive force
\cite{Lum1999,Meyer2006,Hammer2010}. Measurements show that the
hydrophobic force decays exponentially with a decay length on the order of 1 nm
\cite{Israelachvili1984, Marcelja1977,Christenson2001,Lin2005}. 
Additionally, the interaction is not pairwise additive, meaning that
the force between any two hydrophobic objects is altered by the presence
of a third object, hydrophobic or otherwise \cite{SilveraBatista1242477}.

\begin{wrapfigure}[10]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{ \footnotesize
A collection of rigid particles: $P_1,$ $P_2,$ $P_3, \ldots$. The
exterior domain $\Omega$ represents the solvent.}
  \label{fig:domain}
\end{wrapfigure}
Recently, PIs RR and YNY developed a mathematical model, called the
hydrophobic attraction potential (HAP) model \cite{Fu2018_SIAM}, that is based
on the physical origin of hydrophobicity. This model addresses the major
shortcomings of molecular dynamics (MD) and continuum approaches. Based
on preliminary results (\S\ref{sec:preliminary_work}), the PIs propose to
extend this HAP model to offer an alternative modeling methodology that
leads to new mathematical ideas and is both physically accurate and
computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles. These results will facilitate optimal design of smart materials by
tuning the geometry and properties of the amphiphilic particles.

\subsection{Hydrophobic Attraction Potential (HAP)}
\label{sec:HAP}
%Based on the physical origin of hydrophobicity,  we have devised a
%functional, called the hydrophobic attraction potential (HAP), to model
%hydrophobic forces~\cite{Fu2018_SIAM}. 
The motivation for the HAP concept stemmed from PI RR's work concerning
energy barriers in membrane
fusion~\cite{RyKlYaCo16,Chetal16}. By applying a mathematical, squared-gradient
theory for hydrophobic attraction between planar
surfaces~\cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977}, PI RR
and collaborators resolved the long-standing issue of accounting for the
energy of a monolayer fissure surface during topological transitions.
Based on PI RR and YNY's coarse-grained membrane modeling
work~\cite{Fu2017}, the investigators devised a gradient theory for
arbitrary collections of hydrophobic and
amphiphilic particles. As a new method, HAP eliminates the costly calculation
of water by treating the solvent implicitly, is capable of representing
biologically-relevant morphologies, and avoids complicated
re-meshing schemes of continuum approaches by utilizing a particle-based
representation.

@@

The above considerations motivate the following definition for HAP:
\begin{align}
\label{HAP}
  \Phi = \gamma \int_{\Omega} \rho |\nabla u|^2 + \rho^{-1}u^2 \,\dif V. 
\end{align}
The integrand contains
the decay length $\rho$
and a dimensionless scalar function $u(\mathbf{x})$ called the water activity.
The integrand has units of an inverse length, so
the volume integral has units of area. Multiplication by interfacial tension
$\gamma$ makes $\Phi$ an energy.

To establish an attraction between interfaces, the water activity is not
arbitrary but rather is the solution of the screened Laplace
equation boundary value problem 
\begin{equation}
  \label{SL}
  -\rho^2 \Delta u + u = 0, \mbox{ } \mathbf{x} \in \Omega, \qquad
  u = f,  \mbox{ } \mathbf{x} \in \partial \Omega, \qquad 
  u(\mathbf{x}) \to 0, \mbox{ as } |\mathbf{x}| \to \infty.
\end{equation}
The boundary values $f$ define the degree of hydrophobicity of the
water-particle interface: $f=1$ describes a hydrophobic interface, and
$f=0$ describes a hydrophilic interface. Thus, $f$ encodes information
about the particles, and the parameters $\rho$ and $\gamma$ encode
information about the quality of the solvent~\cite{Israelachvili1954,
Discher2002}. In practice, we integrate~\eqref{HAP} by parts and
use~\eqref{SL} to obtain
\begin{equation}
\label{eq:HAP_easy}
\Phi = -\gamma \int_{\partial \Omega} \rho u \nabla u \cdot \boldsymbol{\nu} \,\dif S,
\end{equation}
thereby avoiding volume integral calculations.

The equations~\eqref{HAP} and~\eqref{SL} possess a number of
mathematical properties that mirror the phenomenological characteristics
of hydrophobic attraction. Specifically, solutions of~\eqref{SL} yield
an attractive force between hydrophobic bodies separated by water, and
this attraction decreases exponentially with the distance of
separation~\cite{Eriksson1989}. Using boundary layer analysis, the HAP
converges to a surface energy in the zero-decay length
limit~\cite{Lee2018, Lin2015, Shibata2004}. Finally, we have
demonstrated that the forces derived from HAP theory are
non-additive~\cite{Meyer2006, Fu2018_SIAM}. 
%The hydrophobic force has been implicated in the directed folding of proteins, 
%adhesion between biological membranes, but it is still unanswered as to whether
%the hydrophobic force of the form (\ref{HAP}--\ref{SL}) exists between small molecules.

@@

To define the particle dynamics, let $P_1$, $P_2$, \ldots, $P_N$ be a
finite collection of disjoint, rigid, and closed particles each with
a Lipschitz boundary. Assume that the label $f$ is an element of the

@@

overlap. The total potential, force, and torque are then

@@

To supply viscous dissipation, we incorporate the mobility problem
flow for a rigid body suspension in Stokes flow:
\begin{equation}
\label{eq:stokes}
\begin{aligned}
  &-\mu \Delta \mathbf{u} + \nabla p = 0, \quad \mathbf{x} \in \Omega, \qquad 
  \nabla \cdot \mathbf{u} = 0,  \quad \mathbf{x} \in \Omega,\\
  &{\bf u}(\mathbf{x}) \to 0 \quad \text{as}\ |\mathbf{x}|\to \infty,\qquad 
  \mathbf{u}(\mathbf{x})|_{\partial P_i} = \mathbf{v}_i +
\boldsymbol{\omega}_i\times(\mathbf{x} - \mathbf{a}_i),\\
&\int_{\partial P_i}\boldsymbol{\sigma}\cdot {\bf n} \dif S=-{\bf F}_i, \quad
\int_{\partial P_i} (\mathbf{x} - \mathbf{a}_i)\times (\boldsymbol{\sigma} \cdot \mathbf{n}) \dif S =-{\bf G}_i.
\end{aligned}
\end{equation}
Here, $\mu$ is the fluid viscosity; the first two equations state that
the fluid motion is a divergence-free Stokes flow; the third equation
specifies that the fluid velocity vanishes at infinity; the fourth
equation enforces a rigid body motion on each particle, where
$\mathbf{v}_i$ and $\boldsymbol{\omega}_i$ are unknown translation and
angular velocities; and the last two equations state that the net
viscous forces and torques with fluid shear stress $\boldsymbol{\sigma}$
balance \eqref{eq:total_poten}.

The time integration of
particle configurations goes as follows: \textbf{(i)} solve the
BVP~\eqref{SL} for the screened Laplace equation, \textbf{(ii)}
determine the rigid body forces and torques~\eqref{forceandtorque},
\textbf{(iii)} solve the Stokes mobility problem~\eqref{eq:stokes} for
rigid body motion, and \textbf{(iv)} update the particle configuration.
Numerical challenges and how these challenges are overcome are described
in \S\ref{subsec:specific_aim_2}.

@@

It is worth emphasizing that the HAP model uses only a few parameters:
interfacial tension, decay length, repulsion strength, and particle
shape. For example, an elastic modulus for stretching a vesicle from
micropipette manipulation directly calibrates our interfacial tension
parameter. This is in direct contrast with pair-potential-based
approaches in MD simulations and coarse-grained models where many more
parameters are required~\cite{Varilly2011, Wang2013}.
%MD simulators have also made measurements and lately there is better and better agreement with reality. 
%But even the simplest coarse grained models based on pair potentials for lipids has many more parameters \cite{Varilly2011,Wang2013} .

As a proof of concept, our work has already tested for elastic energies
for bending, stretching, and tilt of the bilayer assembly. The elastic
coefficients derived from the HAP simulations show strikingly positive
agreement with experimentally determined values~\cite{Fu2018_SIAM}.
Encouraged by these results and the hydrodynamic simulations of
bilayer assembly, the PIs propose to extend the HAP model to make direct
comparisons with experimental results \S\ref{subsec:specific_aim_1} and to
establish that the HAP model has the capability to model
flow phenomena across length scales and time scales
\S\ref{subsec:specific_aim_2} using efficient, high-order numerical
methods \S\ref{subsec:specific_aim_3}.

\begin{wrapfigure}[14]{l}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/PW_fig2.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:demixing} \footnotesize An initial assembly of
  small and large particles spontaneously segregates into two smaller
  bodies.}
\end{wrapfigure}
Over the past decades, researchers have used a number of mathematical
tools to simulate vesicles in a shear flow, including lattice
Boltzmann~\cite{KaouiHartingMisbah2011_PRE}, coarse-grained Brownian
dynamics~\cite{NoguchiTakasu2002_BJ}, phase
field~\cite{DuLiuWang2004_JCP,BibenKassnerMisbah2005_PRE}, level
set~\cite{DoyeuxGuyotChabannesEtAl2013_JCAM}, boundary
integral~\cite{Shravan09,Rahimian15}, and immersed boundary
approaches~\cite{KimLai2010_JCP,KimLai2012_PRE,HuLaiSeolEtAl2016_JCP}.
Most of these approaches assume a mathematical surface, whether
implicitly or explicitly, and define an elastic bending energy of the
surface. These vesicle studies built off of numerical methods for
calculating energy minimizing steady equilibrium shapes of lipid bilayer
membranes, vesicles, and red blood cells. These approaches range from
the finite element~\cite{Bartels,Peng13,RyKlYaCo16,Sinha15},
phase field~\cite{Du05,QiangDu08,Lowengrub13}, and immersed
boundary methods~\cite{Hu,Hu13, KimLai2010_JCP}. PI RR and collaborators
led in part the development of phase field functionals of membrane
elastic energy and approaches to coupling membrane elasticity to
fluids~\cite{0951-7715-18-3-016,Du05,DuEuler,QiangDu09}.

@@

Our HAP approach differs from these prior methods in a number of
respects. First, we do not assume a surface. Rather, we 
assume a collection of amphiphilic particles. The
collection of amphiphiles minimize hydrophobic interactions by
sequestering hydrophobic tails in the form of a bilayer, and the
particles' excess free energy gives rise to an elastic bilayer energy.
The second difference lies in the fluid-interface coupling. Here, the
associated mobility problem~\eqref{eq:stokes} is more complicated than
dealing with a stress boundary condition or diffusive surface force
because the fluid velocities are for individual rigid body motions at each
particle surfaces. Finally, the HAP model directly addresses the
existence of multiple phases. We can vary lipid length, spontaneous
curvature, and bending rigidity by introducing different particle shapes
and hydrophobic boundary conditions (Figure~\ref{fig:demixing}). In
contrast, continuum theory deals with multiple phases through additional
surface densities that must satisfy specialized transport
equations~\cite{Lowengrub07, MikuckiZhou17}.

\begin{wrapfigure}[6]{r}{0.35\textwidth}
\centerline{\includegraphics[width=0.35\textwidth]{figures/Lysis.pdf}}
  \vspace{-8pt}
  \caption{\label{fig:lysis} \footnotesize Two hydrophobic particles
  enter and then lyse the circular bilayer.}
\end{wrapfigure}
The greatest strength of using the HAP to model a lipid bilayer membrane is
the ability to form discontinuities (interfacial singularities) from first
physical principles without any artificial manipulation to rearrange the
interface (Figures~\ref{fig:demixing} and~\ref{fig:lysis}). For
example, using the HAP model to simulate a vesicle under a shear flow
that is strong enough to cause the vesicle membrane to rupture (see
\S\ref{sec:preliminary_work}), we expect the HAP model to capture the
reorganization of lipid molecules on the scales of membrane thickness
($\sim 5$ nm), which is a nearly impossible task using phase-field or
immersed boundary methods without extreme refinement around the membrane
and artificial treatment of reconnection during the topological change
of an interface \cite{doi:10.1063/5.0009734, LiAn-Chang16,
doi:10.1098/rspa.2012.0505, doi:10.1137/130941432, Feetzl18,
doi:10.1137/16M1108406}.
